DSSynth: An Automated Digital Controller Synthesis Tool for Physical Plants

### 2017-08-18 ase2017/paper.tex
nudge
improved step response graphs
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

\documentclass[10pt,conference]{IEEEtran}

@@

\renewcommand{\baselinestretch}{0.986}
\newcommand\tool{{DSSynth Toolbox}\xspace}
\newcommand{\fwl}[1]{\mathcal{FWL}[#1]}

@@

We use two different instantiations of the synthesiser. 
The two approaches differ in
the manner in which they check the safety specification.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htb]
{\scriptsize
\centering
\resizebox{.5\textwidth}{!}
{
  \begin{tikzpicture}[scale=0.3,->,>=stealth',shorten >=.2pt,auto, semithick, initial text=, ampersand replacement=\&,]
  \matrix[nodes={draw, fill=none, shape=rectangle, minimum height=.2cm, minimum width=.2cm, align=center
},
          row sep=.6cm, column sep=.9cm] {
   \coordinate (aux1);
   \& \coordinate (aux2);
   \&;\\
   \coordinate (aux3);
   \& \coordinate (aux4);
   \&;\\
   \coordinate (aux5);
   \& \coordinate (aux6);
   \&;\\
   \node[minimum width=1.5cm, minimum height=0.6cm, fill=gray!20] (synth) {{\sc Synthesize}};
   \&
   complexnode/.pic={ 
     \node[rectangle,draw,dotted,
	minimum width=6cm,
	minimum height=1cm,
        pattern=north west lines, pattern color=gray!20,
	label={\sc ~~~~~~~~~~~~Verify},] (verif) {};
     \node[minimum width=1cm, minimum height=0.6cm, fill=gray!20] (verif1) at ([xshift=-2cm]verif.center) {{\sc 1.Safety}};
     \node[minimum width=1cm, minimum height=0.6cm, fill=gray!20] (verif2) at ([xshift=0cm]verif.center) {{\sc 2.Precision}};
     \node[minimum width=1cm, minimum height=0.6cm, fill=gray!20] (verif3) at ([xshift=2cm]verif.center) {{\sc 3.Complete}};
     %\node[minimum width=1cm, minimum height=0.6cm, fill=gray!20] (verif4) at ([xshift=3.1cm]verif.center) {{\sc 5.Sampling}};
   } 
   \& \node[ellipse, fill=gray!20] (done) {{\sc Done}};\\
   \& \\
   \node[minimum height=0cm] (gp) {\sf Program Search};
   \&
   complexnode/.pic={ 
     \coordinate (aux);
   \node (bmc) at ([xshift=-2cm]aux.center) {\sf BMC-based \\ \sf Verifier};
   \node (fp)  at ([xshift=0cm]aux.center) {\sf Fixed-point \\ \sf Arithmetic\\ \sf Verifier};
   \node (sv)  at ([xshift=2cm]aux.center) {\sf Completeness\\ \sf Verifier};
   %\node (cv)  at ([xshift=3.1cm]aux.center) {\sf Sampling\\Verifier};
   }   
    \\
  };

@@

 \end{tikzpicture}
}}
\caption{CEGIS with multi-staged verification}
\label{fig:CEGIS-precision-increment}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The first approach, presented in Fig.~\ref{fig:CEGIS-precision-increment},
starts by devising a digital controller that stabilizes the model while
remaining safe for a pre-selected time horizon ($k$) and a single initial
state; then, it employs a multi-staged verification process as follows:
(i)~The first verification stage ({\sc safety}) checks that the candidate
solution, which we synthesized to be safe for at least one initial
state, is safe for \emph{all} possible initial states, i.e., does not reach
an unsafe state within $k$ steps.
(ii)~The second verification stage ({\sc precision})
 restores soundness with respect to the plant precision
by using interval arithmetic~\cite{moore1966interval} to validate the 
operations performed by the previous stage. 
(iii)~The third verification stage ({\sc complete}) checks that the current
$k$ is large enough to ensure safety for any $k'{>}k$.  Here, we compute the
completeness threshold $\overline{k}$ for the current candidate controller 
$K$~\cite{abatecav2017}, i.e., the number of
iterations required to sufficiently unwind the closed-loop state-space
model such that the boundaries are not violated for any larger number of
iterations, and
check that $k{\geq}\overline{k}$.
%
\begin{figure}
\centering
{\scriptsize
  \resizebox{.5\textwidth}{!}
  {
\begin{tikzpicture}[scale=0.3,->,>=stealth',shorten >=.2pt,auto, semithick, ampersand replacement=\&,]
  \matrix[nodes={draw, fill=none, shape=rectangle, minimum height=.2cm, minimum width=.2cm, align=center},row sep=.8cm, column sep=.2cm] {
   \coordinate (aux1);
   \& \coordinate (aux2);
   \& ;\\
   \&\node[fill=gray!20,align=center,xshift=-1.5cm] (pre) {{\sc 1. pre-}\\{\sc processing}};
   \& \node[fill=gray!20,align=center] (abstract) {\sc 4. abstract};
   \& \coordinate (aux); \\ 
   \&
   \& \node[fill=gray!20,align=center] (synth) {\sc 2. synthesize};
   \& \node[fill=gray!20,align=center, minimum width=4cm] (verify) {\sc 3. verify ($\phi$)};
      \node[draw=none] (SAT) at ([xshift=2.7cm,yshift=-.3cm]verify)  {\sc PASS};
   \& \node[ellipse, fill=gray!20] (done) {{\sc Done}};\\
   \&
   \& \node[draw,rectangle,align=center] (KSAT) { \sf Program \\ \sf Search};
   \& complexnode/.pic={
     \coordinate (AA);
     \node[draw,rectangle,align=center] (AAV) at ([xshift=-1.5cm]AA.center) { \sf Abstract\\ \sf Acceleration};
     \node[draw,rectangle,align=center] (AAC) at ([xshift=1.5cm]AA.center) { \sf Abstraction \\ \sf Verifier};
    }\\
  };
  \path
    (pre.east) edge node[align=center] {} (abstract.west)
    (abstract.south) edge node{} (synth.north)
    ([yshift=1em]synth.east) edge node {$K$} ([yshift=1em]verify.west)
    (aux) edge (abstract.east) 
    (verify.east) edge (done.west);
  \path
    ([xshift=+.6cm]KSAT.north) edge[left] node[xshift=-.3cm] {} ([xshift=0.6cm]synth.south) 
    ([xshift=-.6cm]synth.south) edge node[align=center,xshift=.3cm] {} ([xshift=-.6cm]KSAT.north)
    ([yshift=.5cm]AAV.east) edge node{} ([yshift=.5cm]AAC.west)
    ([yshift=-.5cm]AAC.west) edge node{} ([yshift=-.5cm]AAV.east)
    (AAC.north) edge node{} ([xshift=4.9cm]verify.south)
    ([xshift=-5cm]verify.south) edge node{} (AAV.north);
  \path[-] 
     (verify.north) edge node {C-ex} (aux);
\end{tikzpicture}
}
}
\caption{Abstraction-based CEGIS}
\label{fig:CEGARIS} 
\end{figure}

Instead of unfolding the dynamics, the second approach (illustrated in
Fig.~\ref{fig:CEGARIS}) employs {\em abstract
acceleration}~\cite{cattaruzza2015unbounded} to simultaneously evaluate all
possible progressions of the model.  This eliminates the need for a
completeness threshold, since the time horizon considered in abstract
acceleration is infinite.  An~additional preprocessing phase is used to
compute a set of initial bounds on $K$ based on input constraints.  Note
that these bounds will be used by the {\sc synthesize} phase to reduce the
size of the solution space.
%
Since this model is based on abstraction, we initially perform the synthesis
over a basic model that abstracts away the complexity of the specification
and is only refined to a more complex model when a counterexample is found.
%
% Thus, we may only look at the gain of the
% model to synthesise a controller and to add further restrictions on specific sections of the reach
% tube when these prove to invalidate the solution.
%
The use of this Counterexample-Guided Abstraction-Refinement results in much
faster synthesis and verification since it is applied to both phases,
allowing for faster results whilst maintaining soundness for an infinite
time horizon.

@@

%---------------------------------------------------
\subsubsection{Command Line Version}
%---------------------------------------------------

@@

%---------------------------------------------------
\subsubsection{MATLAB Application Version} 
%---------------------------------------------------

@@

For all evaluated benchmarks, the signal input range related to the control
system is between $-1$ and $1$.  The implementation features used for
digital controller are: $8$~bits for the integer part (including the sign
bit) and $8$~bits for the fractional part.  We~have run the \tool using
CEGIS with multi-staged verification (cf.~Section~\ref{sec:CEGIS}).  All
experiments with \tool v$1$.$0$.$0$ were conducted on an otherwise idle
Intel Core i$7-2600$ $3.40$ GHz processor, with $24$ GB of RAM, running
Ubuntu $64$-bit.

@@

Regarding the state-space representation, we have reproduced the stability property for 
the closed-loop control system of a pendulum represented by Eq.~\eqref{ss-synth-controller}. 
%
\begin{equation}
\label{ss-synth-controller}
\begin{split}
x(n+1) &= A x(n) + B u(n)
\\
y(n) &= C x(n) + D u(n), 
\end{split}\
\end{equation}
%
\noindent where
%
$$A = \left[\begin{array}{cc}-1.999909737361046&-1.000000000000000\\1&0\end{array}\right],$$
$$B = \left [\begin{array}{c}4\\0\end{array}\right],$$
$$C = \left [\begin{array}{cc}-1.756887232846049&-1.7494556607764090\end{array}\right],$$
$$\quad D = \left [\begin{array}{c}9.8\end{array}\right].$$

For this particular closed-loop control system, the feedback matrix (see
Eq.~\eqref{ss-synth-feedback}) generated by the \tool also leads to a stable
system, as illustrated in Fig.~\ref{ss-step-response}.
%
\begin{equation}
\label{ss-synth-feedback}
\begin{split}
K &= \left [\begin{array}{cc}-0.5898&-0.1914\end{array}\right]
\end{split}\
\end{equation}

@@

%The digital controllers synthesized by \tool are all stable using the MATLAB step response
%as a validation procedure, which means that our results are reliable for control engineers. 
The synthesized controllers are also confirmed to be safe outside of our model
representation using MATLAB, which means that our results are usable for control engineers. 
In~particular, given that the closed loop dynamics are convergent, 
the controller's safety can be checked by computing the initial state's infinity norm, i.e., 
the product between $l_{\infty}$ norm of initial states ($\vert\vert{x_{0}}\vert\vert_{\infty}=\max{x_{0}}$) 
and the $l_{1}$ norm of ($\vert\vert{\bar{A}}\vert\vert_{1}=\max\limits_{j}{\sum\limits_{i=1}^{m} \vert \bar{a}_{ij} \vert}$, 
where $\bar{A}=A-BK$). This is a conservative safety check, which is based on a sufficient condition~\cite{toscano2013structured}. 
An~alternative (but not conservative) approach is to unfold the system dynamics up to a given depth, 
with the initial state equal to its maximum value, compute the system input and output via the feedback matrix, 
and then check the state bounds.

@@

The \tool can automatically synthesize stable and safe digital controllers
for intricate dynamic physical plants, represented in MATLAB either as a
transfer function or a state-space equation.  The \tool is the first fully
automated synthesis tool that is algorithmically and numerically sound,
considering various error sources in the implementation of the digital
control algorithm and in the computational modeling of plant
dynamics~\cite{abate2017, abatecav2017}.
%
The \tool presents a practical application of CEGIS-based program synthesis.
Our results in control engineering encourage the application
of this technique in other fields such as program repair or superoptimisation.
%
Additionally, given the current literature in controller synthesis, there is
no other MATLAB toolbox for synthesizing stable and safe digital controllers
for physical plants that considers all aspects of their digital
implementation.
%
As future work, the \tool will perform synthesis considering performance
requirements and will be combined with the
DSVerifier/DSValidator Toolboxes~\cite{issta2017,dsvalidator} to
verify/validate other controller properties.
We will also pursue the application of CEGIS to further software
engineering problems.

>>>>

\documentclass[conference]{IEEEtran}

@@

\renewcommand{\baselinestretch}{0.986}
\setlength{\belowcaptionskip}{-2.8 pt} 
\newcommand\tool{{DSSynth Toolbox}\xspace}
\newcommand{\fwl}[1]{\mathcal{FWL}[#1]}

@@

We use two different instantiations of the synthesiser. 
The two approaches differ in
the manner in which they check the safety specification.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htb]
{\scriptsize
\centering
\resizebox{.5\textwidth}{!}
{
  \begin{tikzpicture}[scale=0.3,->,>=stealth',shorten >=.2pt,auto, semithick, initial text=, ampersand replacement=\&,]
  \matrix[nodes={draw, fill=none, shape=rectangle, minimum height=.2cm, minimum width=.2cm, align=center
},
          row sep=.6cm, column sep=.9cm] {
   \coordinate (aux1);
   \& \coordinate (aux2);
   \&;\\
   \coordinate (aux3);
   \& \coordinate (aux4);
   \&;\\
   \coordinate (aux5);
   \& \coordinate (aux6);
   \&;\\
   \node[minimum width=1.5cm, minimum height=0.6cm, fill=gray!20] (synth) {{\sc Synthesize}};
   \&
   complexnode/.pic={ 
     \node[rectangle,draw,dotted,
	minimum width=6cm,
	minimum height=1cm,
        pattern=north west lines, pattern color=gray!20,
	label={\sc ~~~~~~~~~~~~Verify},] (verif) {};
     \node[minimum width=1cm, minimum height=0.6cm, fill=gray!20] (verif1) at ([xshift=-2cm]verif.center) {{\sc 1.Safety}};
     \node[minimum width=1cm, minimum height=0.6cm, fill=gray!20] (verif2) at ([xshift=0cm]verif.center) {{\sc 2.Precision}};
     \node[minimum width=1cm, minimum height=0.6cm, fill=gray!20] (verif3) at ([xshift=2cm]verif.center) {{\sc 3.Complete}};
     %\node[minimum width=1cm, minimum height=0.6cm, fill=gray!20] (verif4) at ([xshift=3.1cm]verif.center) {{\sc 5.Sampling}};
   } 
   \& \node[ellipse, fill=gray!20] (done) {{\sc Done}};\\
   \& \\
   \node[minimum height=0cm] (gp) {\sf Program Search};
   \&
   complexnode/.pic={ 
     \coordinate (aux);
   \node (bmc) at ([xshift=-2cm]aux.center) {\sf BMC-based \\ \sf Verifier};
   \node (fp)  at ([xshift=0cm]aux.center) {\sf Fixed-point \\ \sf Arithmetic\\ \sf Verifier};
   \node (sv)  at ([xshift=2cm]aux.center) {\sf Completeness\\ \sf Verifier};
   %\node (cv)  at ([xshift=3.1cm]aux.center) {\sf Sampling\\Verifier};
   }   
    \\
  };

@@

 \end{tikzpicture}
}}
\caption{CEGIS with multi-staged verification}
\label{fig:CEGIS-precision-increment}
\end{figure}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
The first approach, presented in Fig.~\ref{fig:CEGIS-precision-increment},
starts by devising a digital controller that stabilizes the model while
remaining safe for a pre-selected time horizon ($k$) and a single initial
state; then, it employs a multi-staged verification process as follows:
(i)~The first verification stage ({\sc safety}) checks that the candidate
solution, which we synthesized to be safe for at least one initial
state, is safe for \emph{all} possible initial states, i.e., does not reach
an unsafe state within $k$ steps.
(ii)~The second verification stage ({\sc precision})
 restores soundness with respect to the plant precision
by using interval arithmetic~\cite{moore1966interval} to validate the 
operations performed by the previous stage. 
(iii)~The third verification stage ({\sc complete}) checks that the current
$k$ is large enough to ensure safety for any $k'{>}k$.  Here, we compute the
completeness threshold $\overline{k}$ for the current candidate controller 
$K$~\cite{abatecav2017}, i.e., the number of
iterations required to sufficiently unwind the closed-loop state-space
model such that the boundaries are not violated for any larger number of
iterations, and
check that $k{\geq}\overline{k}$.
%
\begin{figure}
\centering
{\scriptsize
  \resizebox{.5\textwidth}{!}
  {
\begin{tikzpicture}[scale=0.3,->,>=stealth',shorten >=.2pt,auto, semithick, ampersand replacement=\&,]
  \matrix[nodes={draw, fill=none, shape=rectangle, minimum height=.2cm, minimum width=.2cm, align=center},row sep=.8cm, column sep=.2cm] {
   \coordinate (aux1);
   \& \coordinate (aux2);
   \& ;\\
   \&\node[fill=gray!20,align=center,xshift=-1.5cm] (pre) {{\sc 1. pre-}\\{\sc processing}};
   \& \node[fill=gray!20,align=center] (abstract) {\sc 4. abstract};
   \& \coordinate (aux); \\ 
   \&
   \& \node[fill=gray!20,align=center] (synth) {\sc 2. synthesize};
   \& \node[fill=gray!20,align=center, minimum width=4cm] (verify) {\sc 3. verify ($\phi$)};
      \node[draw=none] (SAT) at ([xshift=2.7cm,yshift=-.3cm]verify)  {\sc PASS};
   \& \node[ellipse, fill=gray!20] (done) {{\sc Done}};\\
   \&
   \& \node[draw,rectangle,align=center] (KSAT) { \sf Program \\ \sf Search};
   \& complexnode/.pic={
     \coordinate (AA);
     \node[draw,rectangle,align=center] (AAV) at ([xshift=-1.5cm]AA.center) { \sf Abstract\\ \sf Acceleration};
     \node[draw,rectangle,align=center] (AAC) at ([xshift=1.5cm]AA.center) { \sf Abstraction \\ \sf Verifier};
    }\\
  };
  \path
    (pre.east) edge node[align=center] {} (abstract.west)
    (abstract.south) edge node{} (synth.north)
    ([yshift=1em]synth.east) edge node {$K$} ([yshift=1em]verify.west)
    (aux) edge (abstract.east) 
    (verify.east) edge (done.west);
  \path
    ([xshift=+.6cm]KSAT.north) edge[left] node[xshift=-.3cm] {} ([xshift=0.6cm]synth.south) 
    ([xshift=-.6cm]synth.south) edge node[align=center,xshift=.3cm] {} ([xshift=-.6cm]KSAT.north)
    ([yshift=.5cm]AAV.east) edge node{} ([yshift=.5cm]AAC.west)
    ([yshift=-.5cm]AAC.west) edge node{} ([yshift=-.5cm]AAV.east)
    (AAC.north) edge node{} ([xshift=4.9cm]verify.south)
    ([xshift=-5cm]verify.south) edge node{} (AAV.north);
  \path[-] 
     (verify.north) edge node {C-ex} (aux);
\end{tikzpicture}
}
}
\caption{Abstraction-based CEGIS}
\label{fig:CEGARIS} 
\end{figure}

The second approach (illustrated in
Fig.~\ref{fig:CEGARIS}) employs {\em abstract
acceleration}~\cite{cattaruzza2015unbounded} to simultaneously evaluate all
possible progressions of the model.  This eliminates the need for a
completeness threshold, since the time horizon considered in abstract
acceleration is infinite.  An~additional preprocessing phase is used to
compute a set of initial bounds on $K$ based on input constraints.  Note
that these bounds will be used by the {\sc synthesize} phase to reduce the
size of the solution space.
%
Since this model is based on abstraction, we initially perform the synthesis
over a basic model that abstracts away the complexity of the specification
and is only refined to a more complex model when a counterexample is found.
%
% Thus, we may only look at the gain of the
% model to synthesise a controller and to add further restrictions on specific sections of the reach
% tube when these prove to invalidate the solution.
%
The use of this Counterexample-Guided Abstraction-Refinement results in much
faster synthesis and verification since it is applied to both phases,
allowing for faster results whilst maintaining soundness for an infinite
time horizon.

@@

%---------------------------------------------------
\subsubsection{Command Line}
%---------------------------------------------------

@@

%---------------------------------------------------
\subsubsection{MATLAB Application} 
%---------------------------------------------------

@@

For all evaluated benchmarks, the signal input range is between $-1$ and $1$.  
The implementation features used for
digital controller are: $8$~bits for the integer part (including the sign
bit) and $8$~bits for the fractional part.  We~have run the \tool using
CEGIS with multi-staged verification (cf.~Section~\ref{sec:CEGIS}).  All
experiments with \tool v$1$.$0$.$0$ were conducted on an otherwise idle
Intel Core i$7-2600$ $3.40$ GHz processor, with $24$ GB of RAM, running
Ubuntu $64$-bit.

@@

Regarding the state-space representation, we have reproduced the stability property for 
the closed-loop control system of a pendulum represented by Eq.~\eqref{ss-synth-controller}. 
%
\begin{equation}
\label{ss-synth-controller}
\begin{array}{c} 
A = \left[\begin{array}{cc}-1.999909737361046&-1.000000000000000\\1&0\end{array}\right], \\
B = \left [\begin{array}{c}4\\0\end{array}\right], \\
C = \left [\begin{array}{cc}-1.756887232846049&-1.7494556607764090\end{array}\right], \\
\quad D = \left [\begin{array}{c}9.8\end{array}\right].
\end{array}
\end{equation}

For this particular closed-loop control system, the feedback matrix (see
Eq.~\eqref{ss-synth-feedback}) generated by the \tool also leads to a stable
system, as illustrated in Fig.~\ref{ss-step-response}.
%
\begin{equation}
\label{ss-synth-feedback}
\begin{split}
K &= \left [\begin{array}{cc}-0.5898&-0.1914\end{array}\right].
\end{split}\
\end{equation}

@@

%The digital controllers synthesized by \tool are all stable using the MATLAB step response
%as a validation procedure, which means that our results are reliable for control engineers. 
The synthesized controllers are also confirmed to be safe outside of our model
representation using MATLAB, which means that our results are usable for control engineers. 
Given that the closed loop dynamics are convergent, 
the controller's safety can be checked by computing the initial state's infinity norm, i.e., 
the product between $l_{\infty}$ norm of initial states ($\vert\vert{x_{0}}\vert\vert_{\infty}=\max{x_{0}}$) 
and the $l_{1}$ norm of ($\vert\vert{\bar{A}}\vert\vert_{1}=\max\limits_{j}{\sum\limits_{i=1}^{m} \vert \bar{a}_{ij} \vert}$, 
where $\bar{A}=A-BK$). This is a conservative safety check, which is based on a sufficient condition~\cite{toscano2013structured}. 
An~alternative (but not conservative) approach is to unfold the system dynamics up to a given depth, 
with the initial state equal to its maximum value, compute the system input and output via the feedback matrix, 
and then check the state bounds.

@@

The \tool can automatically synthesize stable and safe digital controllers
for dynamic physical plants, represented in MATLAB either as a
transfer function or a state-space equation.  The \tool is the first fully
automated synthesis tool that is algorithmically and numerically sound,
considering various error sources in the implementation of the digital
control algorithm and in the computational modeling of plant
dynamics~\cite{abate2017, abatecav2017}.
%
The \tool presents a practical application of CEGIS-based program synthesis.
Our results in control engineering encourage the application
of this technique in other fields such as program repair or superoptimisation.
%
Additionally, given the current literature in controller synthesis, there is
no other MATLAB toolbox for synthesizing stable and safe digital controllers
for physical plants that considers all aspects of their digital
implementation.
%
As future work, the \tool will perform synthesis considering performance
requirements and will be combined with the
DSVerifier/DSValidator Toolboxes~\cite{issta2017,dsvalidator} to
verify/validate other controller properties.
We will also pursue the application of CEGIS to further software
engineering problems.
